
\begin{acknowledgements}
At the end of my undergraduate career, I was convinced pursuing a PhD was not for me. At the time, I did not think I had the drive or intellect to be successful in a graduate program. 
It took the support of many people to not only change my mind, but also get me through the last six years. 
I am so thankful and glad to have them in my life.

My parents, Sue Schadewald and Brad Jackson have always stressed the value of education. They have always supported me in my endeavors, scholastic or otherwise.
My mom is one of the first people I will call for advice.
She has helped me navigate my relationships and growing into being an adult with adult responsibilities.
My dad is the first person I go to for help. He has driven with me across the country twice, first for an internship in Oregon, and second for my program in Berkeley.
I thank my parents for the joy I find in building, fixing, and improving things, my sense of humor, and providing the example of love and selflessness that I strive to emulate.
My younger sister Sadie Jackson has almost beat me to being the first Dr. Jackson, despite being three years younger. 
She is someone I look up to and I thank her for being one of my best friends. 

My new family, my partner Taryn Farber, and our pets Aspen and Bella, have been a constant source of comfort and support.
Having been through her own graduate struggles, Taryn understands and empathizes when I am struggling.
Aspen and Bella do not, but they are always for giving affection and love whether I need it or not.
My childhood best friends Matt Peat and Trent Delehanty remain best friends today, and have offered a lifetime of advice and support.
I thank my college friends Jason LaFave and Claude Christensen for being the family away from home then, and still best friends now.

Finally, I thank my friends and colleagues of Lab11, who, through their work, culture, and camaraderie set the benchmark for what I envisioned a research lab group should be.
When visiting different potential graduate schools and lab groups, Lab11 was the standard by which I judged them. Nowhere else could compare.
I have Josh Adkins to thank for being my brilliant and aloof roommate and co-founder of the Lab11 Berkeley Annex.
It was a rare and wonderful journey being the first of Prabal's Berkeley students, setting up a new lab, doing class projects, and discovering the wonders of the Bay Area. Josh's drive to build things and make a difference kick-started my own enthusiasm.
Will Huang and Jean-Luc Watson have been some of my best friends through graduate school.
I thank Will for his tireless support, as he has undoubtedly helped me more than anyone with my research, writing, and fighting for bounty in the bayou. 
I thank Jean-Luc for keeping lab culture alive with game nights and Utah ``scones.'' 
I also thank Branden Ghena, the unofficial ``Director of Undergraduate Researchers'' for introducing me to the lab and being the most approachable and patient graduate student I have ever met. 
His example led me to be a more patient and joyful teacher. 
I thank Meghan Clark and Pat Pannuto for being stand in advisors as well as friends.
I thank Noah Klugman and Lane Powell for their kindness, honesty, randomly dropping by my apartment, and for being my dog's best friends.
I thank Brad Campbell for the big beers and Deepika Natarajan for being a consistent friend since 470. 
I thank Thomas Zachariah for being my TA at Michigan, as well as a travel buddy in Hong Kong.
Lastly, I thank Andreas Biri, Matt Podolsky, Shishir Patil, Alvin Tan, and Tess Despres for their feedback and support.
Getting through graduate school was only possible with your support. 

I still remember my last October as an undergraduate when Professor Prabal Dutta persuaded me to pursue a PhD.
He knew that I had what it took to be successful, even though I was doubting myself. 
I have him to thank, not only for persuading me to try, but also for his commitment to seeing me through all the way to graduation.
Lab11 set the standard for a creative and productive research group, and Prabal set the standard for a kind, empathetic, and inspiring advisor.
He is someone that will fight for his students, in academia as well as when they get caught on CCTV ``improving'' his faculty portrait in the Cory Hall photo case.
I also thank the other members of my committee: Professors Kris Pister and Stefano Schiavon for their insight and suggestions to make this dissertation the best version I could have created. 
I am a better engineer, researcher, and person having had their guidance.




\vspace*{\fill}
\noindent\small{This work was supported in part by the CONIX Research Center,
one of six centers in JUMP, a Semiconductor Research Corporation
(SRC) program sponsored by DARPA, 
and with support from the
Building Technology Urban Systems group at Lawrence Berkeley
National Laboratory (LBNL). 
Additionally, this material is
based upon work supported by the National Science Foundation
under grant numbers CNS-1824277 and DGE-1106400, and NSF/Intel CPS Security under grant CNS-1822332.
}
\end{acknowledgements}